### 2022-10-22 dissertation/chapters/05-probability.tex
🚧 WIP - Save point (compiles)

Probability theory and statistics can be traced back to as early as the 17th century. In 1718, \citeauthor{ref:demoivre:1718} \cite{ref:demoivre:1718} published the \textit{Doctrine of of Chance}; a book that is widely seen as the first published book on probability theory. In 1763, Thomas Bayes \cite{ref:bayes:1763} published an article titled \textit{An Essay towards solving a Problem in the Doctrine of Chances} where the first version of Bayes' Theorem was introduced. It is from this theorem that the \ac{BHH} was named after.

Probability theory and statistics play a vital role in the \ac{ML} space. A lot of constructs that arise in \ac{ML} originate from the fields of statistics and probability theory. There are many examples of how these two fields overlap. In 1991, \citeauthor{ref:denker:1991}, \cite{ref:denker:1991} proposed a way to transform \ac{ANN} outputs to probability distributions. In 1993,
\citeauthor{ref:neal:1993} \cite{ref:neal:1993} developed the first \ac{MCMC} sampling algorithm for Bayesian \acp{NN}. These are just a  few examples of how the fields of probability theory, statistics and \ac{ML} overlap.

@@

In everyday conversation, the term \textit{probability} is a measure of one's belief in the occurrence of a future event \cite{ref:wackerly:2014}. Probability is a necessary tool used in many fields including physics, biology, chemistry and computer science. These fields contain many cases that generate observations that can not be predicted with absolute certainty \cite{ref:wackerly:2014}. Probability can be be inferred and confirmed through past events. These events are referred to as \textit{random} or \textit{stochastic} events. The probability that a certain event $A$ might occur is denoted $P(A)$. Although these random events cannot be predicted with absolute certainty, the relative frequency with which they occur over many trials if often remarkably stable. Consider flipping an unbiased, fair coin. The coin has two possible outcomes. From this, one can conclude that each side has a $\frac{1}{2}$ or $50\%$ chance of occurring. In statistics, the decimal probability notation is used where $0 <= P(A) <= 1$. Suppose the fair coin is thrown 10 times, one is not guaranteed to observe $0.5$ heads and $0.5$ tails. There is is some probability, although small, that the coin might fall on heads $0/10$ times. The probability of such an event occurring is $0.0009765625$. The \ac{CLT} shows that the normalised sum of events tends toward a normal distribution with a mean value of $0.5$ if the number of events observed $N$, is large \cite{ref:wackerly:2014}. The larger the value of $N$, the higher the confidence of mean probability and relative frequency of the event. The stable long-term relative frequency by which a random event occurs provides an intuitive and meaningful measure of belief that a certain event will occur again at some point in the future \cite{ref:wackerly:2014}. The belief that a random event would occur again sometime in the future, at an expected relative frequency could be useful for decision making. It will later be shown that the proposed \ac{BHH} implements a probabilistic selection mechanism that continuously updates its beliefs such that it learns what the relative frequency by which a heuristic is selected should be.

@@

The occurrence of a given random event $A$ can often be dependent on the occurrence of another event $B$. In the field of medicine, an example of this is to calculate the probability of a certain diagnosis of a sick patient given his/her symptoms. This is referred to as the conditional probability between two events. The conditional probability is expressed as $P(A \vert B)$ and is read \textit{the probability of $A$ given $B$}. On the contrary, the \textit{unconditional} probability is the probability of an event, not dependent on any other. The conditional probability of $A$ given $B$ can be expressed as is given in Definition \ref{eq:probability:cond_probability} below \cite{ref:wackerly:2014}.

@@

Often times probabilities are calculated by considering multiple random events together. Suppose there are two random events $A$ and $B$, then one can calculate the probability of the union and intersection of these events. From this concept, two laws of probability can be formulated. These are referred to as the \textit{multiplicative} and \textit{additive} laws of probability \cite{ref:wackerly:2014} and is given below in Theorems \ref{th:probability:multiple_events:multiplicative} and \ref{th:probability:multiple_events:additive} respectively.

@@

Bayes' Theorem, named after Thomas Bayes, describes the probability of an event $A$, based on prior knowledge of conditions that might be related to $A$ \cite{ref:zalta:2015}. On order to derive the formal theorem and proof, first consider Definition \ref{def:probability:bayes_theorem:partition} below \cite{ref:zalta:2015}.

@@

Probability distributions are mathematical functions that give the probabilities of the occurrences of different possible outcomes in the experiment. They play a crucial role in the architecture and design of the \ac{BHH}. It will later be shown how the proposed \ac{BHH} makes extensive use of probability distributions and sampling to formulate its selection mechanism. The theory and equations presented in the following sections where all taken from \citeauthor{ref:wackerly:2014} \cite{ref:wackerly:2014}.

@@

The Beta distribution is a family of univariate continuous probability distributions over some $x$, with support on the interval $[0,1]$ \cite{ref:wackerly:2014}. It is parameterised by two shape parameters $\alpha > 0$, $\alpha \in \mathbb{R}$ and $\beta > 0$, $\beta \in \mathbb{R}$. The Beta distribution is denoted as $Beta(\alpha, \beta)$. The \ac{PDF} of the Beta distribution is given below in Equation \ref{eq:probability:probability_distributions:beta:pdf}.

@@

\begin{figure}[htbp]
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{images/beta_various.png}
        \caption{Beta distribution}
        \label{fig:probability:probability_distributions:beta_normal}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{images/beta_cumulative.png}
        \caption{Cumulative Beta distribution}
        \label{fig:probability:probability_distributions:beta_cumulative}
    \end{subfigure}
    \par\bigskip
    \caption{An illustration of the Beta distribution (left) \cite{ref:beta:2014} as well as the cumulative Beta distribution (right) \cite{ref:cumulativebeta:2014} for various values of $\alpha$ and $\beta$ .}
    \label{fig:probability:probability_distributions:beta}
\end{figure}

@@

The Dirichlet distribution is a family of multivariate continuous probability distributions over some $x$ in $K$ dimensions  \cite{ref:wackerly:2014}. The Dirichlet distribution is the multivariate generalization of the Beta distribution and is thus sometimes referred to by its alternative name, the Multivariate Beta Distribution. The Dirichlet distribution is parameterised by some vector $\alpha = (\alpha_{1}$,  $\dots, \alpha_{K})$ $\forall_{k=1}^{K} \alpha_{k} > 0$, $\alpha_{k} \in \mathbb{R}$. $\alpha$ is referred to as the concentration parameter. It is later shown in Chapter \ref{chap:bhh} that this parameter plays a vital role in the optimisation process of the \ac{BHH}. The Dirichlet distribution of order $K \geq 2$ with parameters $\alpha$, denoted $Dir(\alpha)$, has a \ac{PDF} as given in Equation \ref{eq:probability:probability_distributions:dirichlet:pdf} below.

@@

\begin{figure}[htbp]
    \includegraphics[width=\textwidth]{images/dirichlet.jpg}
    \caption{The \acp{PDF} for the Dirichlet distribution over the 2-simplex. The concentration parameter $\alpha$ is varied. The values of $\alpha$ are set to (1.5, 1.5, 1.5), (5, 5, 5), (1, 2, 2), and (2, 4, 8) respectively, read from top to bottom, left to right. The values of the \ac{PDF} are shown by the color maps with contour lines at equal values as indicated in the color bars \cite{ref:dirichlet:2020}.}
    \label{fig:probability:probability_distributions:dirichlet}
\end{figure}

@@

The Bernoulli distribution is a discrete probability distribution over some random variable $x$ that takes the value of $1$ with probability $\theta$ and $0$ with probability $1-\theta$  \cite{ref:wackerly:2014}. It is denoted as $Ber(\theta)$ with support $x \in \{0, 1\}$. In probability theory, the Bernoulli distribution is often used to explain the possible outcomes of a single experiment that asks a \textit{yes-no} question such as flipping a coin. The outcome of such an experiment is a boolean value. The Bernoulli distribution has a \ac{PMF} as given below in Equation \ref{eq:probability:probability_distributions:bernoulli:pmf}.

@@

The mean of the Bernoulli distribution approaches $\theta$ over many samples according to the \ac{CLT} \cite{ref:grinstead:1997}. Figure \ref{fig:probability:probability_distributions:bernoulli:coin} shows a simulation of a fair-coin flipping simulation. One can see how the mean converges to 0.5 for various sample sizes.

@@

The Binomial distribution is a discrete probability distribution over a random variable $x$ taking on a number of successes in $N$ sequential independent experiments that each ask a \textit{yes-no} question \cite{ref:wackerly:2014}. The probability of a single independent experiment yielding a success is given as $\theta$ and the Binomial distribution is denoted as $Bin(N, \theta)$ with support $x \in \{0, 1, \dots, N\}$.  It should be noted that the Binomial distribution is the extension of the Bernoulli distribution over $N$ independent sequential experiments and thus, each experiment also yields some boolean outcome. When $N=1$, the experiment is referred to as a Bernoulli trial and the distribution is just a Bernoulli distribution. When $N > 1$, the sequence of outcomes is referred to as a Bernoulli process. The \ac{PMF} of the Binomial distribution is given in Equation \ref{eq:probability:probability_distributions:binomial:pmf} below.

@@

The Categorical distribution is a discrete probability distribution over some random variable $x$, taking on any one of $K$ possible categories \cite{ref:wackerly:2014}. There is no innate underlying ordering to these categories and therefore, for simplicity, each category is assigned a numerical representative value such that $k = (1, \dots, K)$. The probabilities for all outcomes is given by the probability vector $\theta = (\theta_{1}, \dots, \theta_{K})$.  This means that the probability $P(x=k)=\theta_{k}$ has support $x \in \{1, \dots, K\}$. The Categorical distribution, denoted $Cat(\theta)$ is the generalization of the Bernoulli distribution and is sometimes referred to it by its alternative names, the generalised Bernoulli Distribution or the Multinoulli distribution. In probability theory, the Categorical distribution is often used to explain the outcome of a single experiment with more than two possible outcomes such as rolling a die \cite{ref:wackerly:2014}. The \ac{PMF} of the Categorical distribution is given in Equation \ref{eq:probability:probability_distributions:categorical:pmf} below.

@@

where $[x = k]$ is the Iversion Bracket \cite{ref:iverson:1962}, yielding 1 if $x = k$ and 0 otherwise. From this, one can conclude that for a given class $k$, the categorical distribution simply yields $\theta_{k}$ as is shown in Equation \ref{eq:probability:probability_distributions:categorical:pmf_k} below.

@@

The Multinomial distribution is a discrete probability distribution over some random variable $x = (x_{1}, \dots\, x_{K})$ that takes on the counts for each occurrence of $K$ possible classes in $N$ independent trials \cite{ref:wackerly:2014}. The probabilities for all possible outcomes in a single trial is given by the probability vector $\theta = (\theta_{1}, \dots, \theta_{K})$. The Multinomial distribution, denoted $Mul(N, K, \theta)$, is thus the generalization of the Binomial distribution to $K$ dimensions. Consider the following special cases.

@@

The conjugate prior to a Categorical and Multinomail distribution is the Dirichlet distribution \cite{ref:wackerly:2014}. Similar to the proof of the conjugate prior for the Bernoulli distribution as shown in Section \ref{eq:probability:conjugate_priors:binom_likelihood:posterior} above, this means that the posterior distribution must have the same functional form $\mathcal{A}(v)$ as the prior distribution. This is shown as follows. \\\\
\textbf{Setup}:

@@

In general, there are two main views to probability and statistics. These include the \textit{frequentist} and the \textit{Bayesian} view of statistics. Naturally, Bayesian statistics is based on Bayes' Theorem as was presented in Section \ref{sec:probability:bayes_theorem}. Bayesian statistics describe the probability of an event in terms of some belief based on previous knowledge of the event and the conditions under which the event happens \cite{ref:hackenberger:2019}. To introduce the concept of Bayesian inference and analysis, one first need to understand the difference between these two approaches. Bayesian statistics out-date the frequentist approach, but lacked interest in the early days partly because of the limited applications where the conjugate priors where known \cite{ref:hackenberger:2019}. More recent advancements in mathematical methods popularised the Bayesian approach again. A notable contribution to this switch was the invention of \ac{MCMC} in the 1950's. This family of algorithms allowed for the construction of random sampling algorithms from a probability distribution which allows for the calculation of Bayesian hierarchical models \cite{ref:hackenberger:2019}. Soon after followed one of the earliest papers that use Bayesian statistics in the field of medicine in 1982 \cite{ref:ashby:2006}.

The difference between the frequentist approach and the Bayesian approach can be illustrated using an example. \citeauthor{ref:hackenberger:2019} \cite{ref:hackenberger:2019} suggests an experiment that investigates whether the sex ration in some hypothetical mice population is $1:1$. Two experiments can be designed. In the first experiment, mice are randomly selected until the first male is chosen. The result in this experiment will then be the total number of mice chosen by sex. For the second experiment, exactly seven mice are randomly selected. The result of the second experiment would be the number of males and females in a sample of seven. Suppose the outcome of the experiment was $FFFFFFM$, where $F$ represents a female and $M$ represents a male. If the experimental design is not known ahead of time, the result is useless. Consider the $P$-value for each of these experiments. The P-value is the probability of obtaining results at least as extreme as the observed results of a statistical hypothesis test \cite{ref:beers:2022}. For the first experiment, the $P$-value is $0.031$ and for the second experiment, the $P$-value is $0.227$. Using a confidence level of $\alpha = 0.05$, one would conclude opposite outcomes for these two experiments when it comes to rejecting the null hypothesis, despite using the same data. The reason for this difference is because of the difference in their null distributions. The first approach uses a geometrical approach and the second used a binomial approach as is illustrated on Figure \ref{fig:probability:bayesian_statistics:mouse_experiment_outcome} below.

\begin{figure}[htbp]
    \includegraphics[width=\textwidth]{images/mouse_experiment_outcome.jpg}
    \caption{The experimental outcomes for the mice-population experiments as was taken from \cite{ref:hackenberger:2019} .}
    \label{fig:probability:bayesian_statistics:mouse_experiment_outcome}
\end{figure}

If Bayesian statistics is used then the experimental design that was chosen does not matter. In Bayesian statistics it is common to use a Beta distribution as a prior distribution. If the prior distribution is sampled from $Beta(3,3)$ then using Bayesian analysis, the posterior distribution, according to the outcomes of this experiment, would yield $Beta(9,4)$. The process by which this is done is shown later in this section. The $Beta$ distribution was presented in Section \ref{sec:probability:probability_distributions:beta}. \citeauthor{ref:hackenberger:2019} \cite{ref:hackenberger:2019} mentions that the $Beta$ distribution can be seen as a probability distribution of the occurrence of specific parameters. From the information that is now known about the $Beta$ distribution, it is possible to calculate the probability that the sex ratio in this mice population is not 1:1 with the $Beta$ distribution as a prior, yielding a $P$-value of $0.92$. This means that there is a probability of $92\%$ that the mice population is not 1:1, regardless of experiment design. This is illustrated in Figure \ref{fig:probability:bayesian_statistics:mouse_distributions} below.

\begin{figure}[htbp]
    \includegraphics[width=\textwidth]{images/mouse_experiment_distributions.jpg}
    \caption{An illustration of the prior and posterior probability distributions for the outcomes of the mice-population experiment, using a $Beta$ prior, as was taken from \cite{ref:hackenberger:2019}.}
    \label{fig:probability:bayesian_statistics:mouse_distributions}
\end{figure}

@@

Bayesian analysis is the process by which prior beliefs are updated as a result of observing new data/evidence. Similar to the approaches followed above to explain Bayesian statistics, a proposal is made to explain Bayesian analysis by means of an example taken from \cite{ref:wackerly:2014}. Let $Y_{1}, Y_{2}, \dots, Y_{N}$ denote the random variable that is observed over a sample size of $N$. Then the likelihood of the sample is given as $\mathcal{L}(y_{1}, y_{2}, \dots, y_{n} \vert \theta)$. In the discrete case this function is defined to be the joint probability $P(Y_{1} = y_{1}, Y_{2} = y_{2}, \dots, Y_{N} = y_{n})$ and the continuous case yields the joint density of $Y_{1}, Y_{2}, \dots, Y_{N}$, evaluated at $y_{1}, y_{2}, \dots, y_{n}$. The Bayesian view models the parameter $\theta$ as a random variable with a probability distribution. This probability distribution is referred to as the prior distribution of $\theta$. The symbol $\theta$ is included in the notation of $\mathcal{L}$ as an argument to illustrate that this function is dependent explicitly on the value of $\theta$. Importantly, this prior distribution is specified before any data is collected and represents the theoretical prior knowledge about $\theta$. Assume that the parameter $\theta$ has a continuous distribution with density $g(\theta)$ that has no unknown parameters. Considering the likelihood of the data and the prior on $\theta$, then the joint likelihood of $Y_{1}, Y_{2}, \dots, Y_{N}, \theta$ is given in Equation \ref{eq:probability:bayesian_statistic:bayesian_analysis:joint_likelihood} below.

@@

\citeauthor{ref:wackerly:2014}\cite{ref:wackerly:2014} mentions that the posterior density summarises all the pertinent information about the parameter $\theta$ by making use of the information contained in the prior for $\theta$ as well as the information in the observed data/evidence. Consider now how Bayesian analysis can used to update the priors on $\theta$ based on newly observed data. As with the example above, the below is taken from \cite{ref:wackerly:2014}.

>>>>

Probability theory and statistics can be traced back to as early as the 17th century. In 1718,~\citeauthor{ref:demoivre:1718}~\cite{ref:demoivre:1718} published the \textit{Doctrine of of Chance}; a book that is widely seen as the first published book on probability theory. In 1763, Thomas Bayes~\cite{ref:bayes:1763} published an article titled \textit{An Essay towards solving a Problem in the Doctrine of Chances} where the first version of Bayes' Theorem was introduced. It is from this theorem that the \ac{BHH} was named after.

Probability theory and statistics play a vital role in the \ac{ML} space. A lot of constructs that arise in \ac{ML} originate from the fields of statistics and probability theory. There are many examples of how these two fields overlap. In 1991,~\citeauthor{ref:denker:1991},~\cite{ref:denker:1991} proposed a way to transform \ac{ANN} outputs to probability distributions. In 1993,
\citeauthor{ref:neal:1993}~\cite{ref:neal:1993} developed the first \ac{MCMC} sampling algorithm for Bayesian \acp{NN}. These are just a  few examples of how the fields of probability theory, statistics and \ac{ML} overlap.

@@

In everyday conversation, the term \textit{probability} is a measure of one's belief in the occurrence of a future event~\cite{ref:wackerly:2014}. Probability is a necessary tool used in many fields including physics, biology, chemistry and computer science. These fields contain many cases that generate observations that can not be predicted with absolute certainty~\cite{ref:wackerly:2014}. Probability can be be inferred and confirmed through past events. These events are referred to as \textit{random} or \textit{stochastic} events. The probability that a certain event $A$ might occur is denoted $P(A)$. Although these random events cannot be predicted with absolute certainty, the relative frequency with which they occur over many trials if often remarkably stable. Consider flipping an unbiased, fair coin. The coin has two possible outcomes. From this, one can conclude that each side has a $\frac{1}{2}$ or $50\%$ chance of occurring. In statistics, the decimal probability notation is used where $0 <= P(A) <= 1$. Suppose the fair coin is thrown 10 times, one is not guaranteed to observe $0.5$ heads and $0.5$ tails. There is is some probability, although small, that the coin might fall on heads $0/10$ times. The probability of such an event occurring is $0.0009765625$. The \ac{CLT} shows that the normalised sum of events tends toward a normal distribution with a mean value of $0.5$ if the number of events observed $N$, is large~\cite{ref:wackerly:2014}. The larger the value of $N$, the higher the confidence of mean probability and relative frequency of the event. The stable long-term relative frequency by which a random event occurs provides an intuitive and meaningful measure of belief that a certain event will occur again at some point in the future~\cite{ref:wackerly:2014}. The belief that a random event would occur again sometime in the future, at an expected relative frequency could be useful for decision making. It will later be shown that the proposed \ac{BHH} implements a probabilistic selection mechanism that continuously updates its beliefs such that it learns what the relative frequency by which a heuristic is selected should be.

@@

The occurrence of a given random event $A$ can often be dependent on the occurrence of another event $B$. In the field of medicine, an example of this is to calculate the probability of a certain diagnosis of a sick patient given his/her symptoms. This is referred to as the conditional probability between two events. The conditional probability is expressed as $P(A \vert B)$ and is read \textit{the probability of $A$ given $B$}. On the contrary, the \textit{unconditional} probability is the probability of an event, not dependent on any other. The conditional probability of $A$ given $B$ can be expressed as is given in Definition \ref{eq:probability:cond_probability} below~\cite{ref:wackerly:2014}.

@@

Often times probabilities are calculated by considering multiple random events together. Suppose there are two random events $A$ and $B$, then one can calculate the probability of the union and intersection of these events. From this concept, two laws of probability can be formulated. These are referred to as the \textit{multiplicative} and \textit{additive} laws of probability~\cite{ref:wackerly:2014} and is given below in Theorems \ref{th:probability:multiple_events:multiplicative} and \ref{th:probability:multiple_events:additive} respectively.

@@

Bayes' Theorem, named after Thomas Bayes, describes the probability of an event $A$, based on prior knowledge of conditions that might be related to $A$~\cite{ref:zalta:2015}. On order to derive the formal theorem and proof, first consider Definition \ref{def:probability:bayes_theorem:partition} below~\cite{ref:zalta:2015}.

@@

Probability distributions are mathematical functions that give the probabilities of the occurrences of different possible outcomes in the experiment. They play a crucial role in the architecture and design of the \ac{BHH}. It will later be shown how the proposed \ac{BHH} makes extensive use of probability distributions and sampling to formulate its selection mechanism. The theory and equations presented in the following sections where all taken from~\citeauthor{ref:wackerly:2014}~\cite{ref:wackerly:2014}.

@@

The Beta distribution is a family of univariate continuous probability distributions over some $x$, with support on the interval $[0,1]$~\cite{ref:wackerly:2014}. It is parameterised by two shape parameters $\alpha > 0$, $\alpha \in \mathbb{R}$ and $\beta > 0$, $\beta \in \mathbb{R}$. The Beta distribution is denoted as $Beta(\alpha, \beta)$. The \ac{PDF} of the Beta distribution is given below in Equation \ref{eq:probability:probability_distributions:beta:pdf}.

@@

\begin{figure}[htbp]
      \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{images/beta_various.png}
            \caption{Beta distribution}
            \label{fig:probability:probability_distributions:beta_normal}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
            \includegraphics[width=\textwidth]{images/beta_cumulative.png}
            \caption{Cumulative Beta distribution}
            \label{fig:probability:probability_distributions:beta_cumulative}
      \end{subfigure}
      \par\bigskip
      \caption{An illustration of the Beta distribution (left)~\cite{ref:beta:2014} as well as the cumulative Beta distribution (right)~\cite{ref:cumulativebeta:2014} for various values of $\alpha$ and $\beta$ .}
      \label{fig:probability:probability_distributions:beta}
\end{figure}

@@

The Dirichlet distribution is a family of multivariate continuous probability distributions over some $x$ in $K$ dimensions ~\cite{ref:wackerly:2014}. The Dirichlet distribution is the multivariate generalization of the Beta distribution and is thus sometimes referred to by its alternative name, the Multivariate Beta Distribution. The Dirichlet distribution is parameterised by some vector $\alpha = (\alpha_{1}$,  $\dots, \alpha_{K})$ $\forall_{k=1}^{K} \alpha_{k} > 0$, $\alpha_{k} \in \mathbb{R}$. $\alpha$ is referred to as the concentration parameter. It is later shown in Chapter \ref{chap:bhh} that this parameter plays a vital role in the optimisation process of the \ac{BHH}. The Dirichlet distribution of order $K \geq 2$ with parameters $\alpha$, denoted $Dir(\alpha)$, has a \ac{PDF} as given in Equation \ref{eq:probability:probability_distributions:dirichlet:pdf} below.

@@

\begin{figure}[htbp]
      \includegraphics[width=\textwidth]{images/dirichlet.jpg}
      \caption{The \acp{PDF} for the Dirichlet distribution over the 2-simplex. The concentration parameter $\alpha$ is varied. The values of $\alpha$ are set to (1.5, 1.5, 1.5), (5, 5, 5), (1, 2, 2), and (2, 4, 8) respectively, read from top to bottom, left to right. The values of the \ac{PDF} are shown by the color maps with contour lines at equal values as indicated in the color bars~\cite{ref:dirichlet:2020}.}
      \label{fig:probability:probability_distributions:dirichlet}
\end{figure}

@@

The Bernoulli distribution is a discrete probability distribution over some random variable $x$ that takes the value of $1$ with probability $\theta$ and $0$ with probability $1-\theta$ ~\cite{ref:wackerly:2014}. It is denoted as $Ber(\theta)$ with support $x \in \{0, 1\}$. In probability theory, the Bernoulli distribution is often used to explain the possible outcomes of a single experiment that asks a \textit{yes-no} question such as flipping a coin. The outcome of such an experiment is a boolean value. The Bernoulli distribution has a \ac{PMF} as given below in Equation \ref{eq:probability:probability_distributions:bernoulli:pmf}.

@@

The mean of the Bernoulli distribution approaches $\theta$ over many samples according to the \ac{CLT}~\cite{ref:grinstead:1997}. Figure \ref{fig:probability:probability_distributions:bernoulli:coin} shows a simulation of a fair-coin flipping simulation. One can see how the mean converges to 0.5 for various sample sizes.

@@

The Binomial distribution is a discrete probability distribution over a random variable $x$ taking on a number of successes in $N$ sequential independent experiments that each ask a \textit{yes-no} question~\cite{ref:wackerly:2014}. The probability of a single independent experiment yielding a success is given as $\theta$ and the Binomial distribution is denoted as $Bin(N, \theta)$ with support $x \in \{0, 1, \dots, N\}$.  It should be noted that the Binomial distribution is the extension of the Bernoulli distribution over $N$ independent sequential experiments and thus, each experiment also yields some boolean outcome. When $N=1$, the experiment is referred to as a Bernoulli trial and the distribution is just a Bernoulli distribution. When $N > 1$, the sequence of outcomes is referred to as a Bernoulli process. The \ac{PMF} of the Binomial distribution is given in Equation \ref{eq:probability:probability_distributions:binomial:pmf} below.

@@

The Categorical distribution is a discrete probability distribution over some random variable $x$, taking on any one of $K$ possible categories~\cite{ref:wackerly:2014}. There is no innate underlying ordering to these categories and therefore, for simplicity, each category is assigned a numerical representative value such that $k = (1, \dots, K)$. The probabilities for all outcomes is given by the probability vector $\theta = (\theta_{1}, \dots, \theta_{K})$.  This means that the probability $P(x=k)=\theta_{k}$ has support $x \in \{1, \dots, K\}$. The Categorical distribution, denoted $Cat(\theta)$ is the generalization of the Bernoulli distribution and is sometimes referred to it by its alternative names, the generalised Bernoulli Distribution or the Multinoulli distribution. In probability theory, the Categorical distribution is often used to explain the outcome of a single experiment with more than two possible outcomes such as rolling a die~\cite{ref:wackerly:2014}. The \ac{PMF} of the Categorical distribution is given in Equation \ref{eq:probability:probability_distributions:categorical:pmf} below.

@@

where $[x = k]$ is the Iversion Bracket~\cite{ref:iverson:1962}, yielding 1 if $x = k$ and 0 otherwise. From this, one can conclude that for a given class $k$, the categorical distribution simply yields $\theta_{k}$ as is shown in Equation \ref{eq:probability:probability_distributions:categorical:pmf_k} below.

@@

The Multinomial distribution is a discrete probability distribution over some random variable $x = (x_{1}, \dots\, x_{K})$ that takes on the counts for each occurrence of $K$ possible classes in $N$ independent trials~\cite{ref:wackerly:2014}. The probabilities for all possible outcomes in a single trial is given by the probability vector $\theta = (\theta_{1}, \dots, \theta_{K})$. The Multinomial distribution, denoted $Mul(N, K, \theta)$, is thus the generalization of the Binomial distribution to $K$ dimensions. Consider the following special cases.

@@

The conjugate prior to a Categorical and Multinomail distribution is the Dirichlet distribution~\cite{ref:wackerly:2014}. Similar to the proof of the conjugate prior for the Bernoulli distribution as shown in Section \ref{eq:probability:conjugate_priors:binom_likelihood:posterior} above, this means that the posterior distribution must have the same functional form $\mathcal{A}(v)$ as the prior distribution. This is shown as follows. \\\\
\textbf{Setup}:

@@

In general, there are two main views to probability and statistics. These include the \textit{frequentist} and the \textit{Bayesian} view of statistics. Naturally, Bayesian statistics is based on Bayes' Theorem as was presented in Section \ref{sec:probability:bayes_theorem}. Bayesian statistics describe the probability of an event in terms of some belief based on previous knowledge of the event and the conditions under which the event happens~\cite{ref:hackenberger:2019}. To introduce the concept of Bayesian inference and analysis, one first need to understand the difference between these two approaches. Bayesian statistics out-date the frequentist approach, but lacked interest in the early days partly because of the limited applications where the conjugate priors where known~\cite{ref:hackenberger:2019}. More recent advancements in mathematical methods popularised the Bayesian approach again. A notable contribution to this switch was the invention of \ac{MCMC} in the 1950's. This family of algorithms allowed for the construction of random sampling algorithms from a probability distribution which allows for the calculation of Bayesian hierarchical models~\cite{ref:hackenberger:2019}. Soon after followed one of the earliest papers that use Bayesian statistics in the field of medicine in 1982~\cite{ref:ashby:2006}.

The difference between the frequentist approach and the Bayesian approach can be illustrated using an example.~\citeauthor{ref:hackenberger:2019}~\cite{ref:hackenberger:2019} suggests an experiment that investigates whether the sex ration in some hypothetical mice population is $1:1$. Two experiments can be designed. In the first experiment, mice are randomly selected until the first male is chosen. The result in this experiment will then be the total number of mice chosen by sex. For the second experiment, exactly seven mice are randomly selected. The result of the second experiment would be the number of males and females in a sample of seven. Suppose the outcome of the experiment was $FFFFFFM$, where $F$ represents a female and $M$ represents a male. If the experimental design is not known ahead of time, the result is useless. Consider the $P$-value for each of these experiments. The P-value is the probability of obtaining results at least as extreme as the observed results of a statistical hypothesis test~\cite{ref:beers:2022}. For the first experiment, the $P$-value is $0.031$ and for the second experiment, the $P$-value is $0.227$. Using a confidence level of $\alpha = 0.05$, one would conclude opposite outcomes for these two experiments when it comes to rejecting the null hypothesis, despite using the same data. The reason for this difference is because of the difference in their null distributions. The first approach uses a geometrical approach and the second used a binomial approach as is illustrated on Figure \ref{fig:probability:bayesian_statistics:mouse_experiment_outcome} below.

\begin{figure}[htbp]
      \includegraphics[width=\textwidth]{images/mouse_experiment_outcome.jpg}
      \caption{The experimental outcomes for the mice-population experiments as was taken from~\cite{ref:hackenberger:2019} .}
      \label{fig:probability:bayesian_statistics:mouse_experiment_outcome}
\end{figure}

If Bayesian statistics is used then the experimental design that was chosen does not matter. In Bayesian statistics it is common to use a Beta distribution as a prior distribution. If the prior distribution is sampled from $Beta(3,3)$ then using Bayesian analysis, the posterior distribution, according to the outcomes of this experiment, would yield $Beta(9,4)$. The process by which this is done is shown later in this section. The $Beta$ distribution was presented in Section \ref{sec:probability:probability_distributions:beta}.~\citeauthor{ref:hackenberger:2019}~\cite{ref:hackenberger:2019} mentions that the $Beta$ distribution can be seen as a probability distribution of the occurrence of specific parameters. From the information that is now known about the $Beta$ distribution, it is possible to calculate the probability that the sex ratio in this mice population is not 1:1 with the $Beta$ distribution as a prior, yielding a $P$-value of $0.92$. This means that there is a probability of $92\%$ that the mice population is not 1:1, regardless of experiment design. This is illustrated in Figure \ref{fig:probability:bayesian_statistics:mouse_distributions} below.

\begin{figure}[htbp]
      \includegraphics[width=\textwidth]{images/mouse_experiment_distributions.jpg}
      \caption{An illustration of the prior and posterior probability distributions for the outcomes of the mice-population experiment, using a $Beta$ prior, as was taken from~\cite{ref:hackenberger:2019}.}
      \label{fig:probability:bayesian_statistics:mouse_distributions}
\end{figure}

@@

Bayesian analysis is the process by which prior beliefs are updated as a result of observing new data/evidence. Similar to the approaches followed above to explain Bayesian statistics, a proposal is made to explain Bayesian analysis by means of an example taken from~\cite{ref:wackerly:2014}. Let $Y_{1}, Y_{2}, \dots, Y_{N}$ denote the random variable that is observed over a sample size of $N$. Then the likelihood of the sample is given as $\mathcal{L}(y_{1}, y_{2}, \dots, y_{n} \vert \theta)$. In the discrete case this function is defined to be the joint probability $P(Y_{1} = y_{1}, Y_{2} = y_{2}, \dots, Y_{N} = y_{n})$ and the continuous case yields the joint density of $Y_{1}, Y_{2}, \dots, Y_{N}$, evaluated at $y_{1}, y_{2}, \dots, y_{n}$. The Bayesian view models the parameter $\theta$ as a random variable with a probability distribution. This probability distribution is referred to as the prior distribution of $\theta$. The symbol $\theta$ is included in the notation of $\mathcal{L}$ as an argument to illustrate that this function is dependent explicitly on the value of $\theta$. Importantly, this prior distribution is specified before any data is collected and represents the theoretical prior knowledge about $\theta$. Assume that the parameter $\theta$ has a continuous distribution with density $g(\theta)$ that has no unknown parameters. Considering the likelihood of the data and the prior on $\theta$, then the joint likelihood of $Y_{1}, Y_{2}, \dots, Y_{N}, \theta$ is given in Equation \ref{eq:probability:bayesian_statistic:bayesian_analysis:joint_likelihood} below.

@@

\citeauthor{ref:wackerly:2014}\cite{ref:wackerly:2014} mentions that the posterior density summarises all the pertinent information about the parameter $\theta$ by making use of the information contained in the prior for $\theta$ as well as the information in the observed data/evidence. Consider now how Bayesian analysis can used to update the priors on $\theta$ based on newly observed data. As with the example above, the below is taken from~\cite{ref:wackerly:2014}.
